---
title: "Linear Regression Review: Example 1"
author: "Samuel Burer (samuel-burer@uiowa.edu)"
date: "Updated: `r format(Sys.time(), '%B %e, %Y')`"
output: html_notebook
---

----

## Preparation

If you are following along in RStudio, to prepare for this example, choose the
RStudio menu item *Session > Restart and Clear Output*. Alternatively, you can
choose *Run > Restart R and Clear Output* in the notebook code window.  You
will also need to install the packages `car` and `ggplot2` and to source
`code_Linear_Regression_Review.R`.

```{r}
# install.packages(c("car", "ggplot2")) # Uncomment and run if necessary
```


```{r}
source("code_Linear_Regression_Review.R")
```

----

## Data

In this example, we will develop a linear-regression model for the sale prices
of single-family homes in the Des Moines area. The data was collected from the
[Polk County Assessor's
website](http://web.assess.co.polk.ia.us/cgi-bin/web/tt/infoqry.cgi?tt=home/index)
and is stored in a data frame called `homes` in the file
`data_Linear_Regression_Review.Rdata`.

```{r}
load("data_Linear_Regression_Review.Rdata")
library(readr)
HR_Exercise_Training <- read_csv("HRExercise.csv", 
    col_types = cols(Date = col_date(format = "%m/%d/%y"), 
        Time = col_time(format = "%T"), DateTime = col_datetime(format = "%m/%d/%y %T"), 
        Exercise = col_factor(levels = c("Yoga", "Walking", "Running", "Biking", "Trail Walking", "Trail Running", "Mountain Biking")), 
        ExerciseValue = col_factor(levels = c("0", "1", "2", "3", "4", "5", "6")), 
        Course = col_factor(levels = c("Alta Vista", 
            "Deer Hollow", "Enterprise South", 
            "Four Pass Loop", "Full Practice", 
            "Greenway", "Hawk's Ridge", "Montreat", 
            "Montreat Trails", "Restorative", 
            "Riverwalk", "Southern Soul", 
            "Stringers", "White Oak - Green", 
            "White Sands Cove", "Whole Enchilada")), 
        CourseValue = col_factor(levels = c("0", "1", "2", "3")),
        Zone = col_factor(levels = c("1", "2", "3", "4", "5")),
        IsYoga = col_factor(levels = c("0", "1")), 
        IsFullPractice = col_factor(levels = c("0", "1")), 
        IsSouthernSoul = col_factor(levels = c("0", "1")), 
        IsOtherYoga = col_factor(levels = c("0", "1")), 
        IsLessThan12HoursSinceLastYoga = col_factor(levels = c("0", "1")), 
        IsLessThan24HoursSinceLastYoga = col_factor(levels = c("0", "1")), 
        IsLessThan36HoursSinceLastYoga = col_factor(levels = c("0", "1")), 
        IsLessThan48HoursSinceLastYoga = col_factor(levels = c("0", "1")), 
        IsLessThan60HoursSinceLastYoga = col_factor(levels = c("0", "1")), 
        IsLessThan72HoursSinceLastYoga = col_factor(levels = c("0", "1")), 
        IsBiking = col_factor(levels = c("0", "1")), 
        IsBikingAtWhiteOakGreen = col_factor(levels = c("0", "1")), 
        IsBikingAtWholeEnchilada = col_factor(levels = c("0", "1")), 
        IsBikingAtStringers = col_factor(levels = c("0", "1")), 
        IsBikingAtRiverwalk = col_factor(levels = c("0", "1")),
        IsBikingAtGreenway = col_factor(levels = c("0", "1")),
        IsOtherBiking = col_factor(levels = c("0", "1")),
        IsLessThan12HoursSinceLastBiking = col_factor(levels = c("0", "1")),
        IsLessThan24HoursSinceLastBiking = col_factor(levels = c("0", "1")),
        IsLessThan36HoursSinceLastBiking = col_factor(levels = c("0", "1")),
        IsLessThan48HoursSinceLastBiking = col_factor(levels = c("0", "1")),
        IsLessThan60HoursSinceLastBiking = col_factor(levels = c("0", "1")),
        IsLessThan72HoursSinceLastBiking = col_factor(levels = c("0", "1")),
        IsWalking = col_factor(levels = c("0", "1")),
        IsWalkingAtHawksRidge = col_factor(levels = c("0", "1")),
        IsWalkingAtFourPassLoop = col_factor(levels = c("0", "1")),
        IsWalkingAtRiverwalk = col_factor(levels = c("0", "1")),
        IsOtherWalking = col_factor(levels = c("0", "1")),
        IsLessThan12HoursSinceLastWalking = col_factor(levels = c("0", "1")),
        IsLessThan24HoursSinceLastWalking = col_factor(levels = c("0", "1")),
        IsLessThan36HoursSinceLastWalking = col_factor(levels = c("0", "1")),
        IsLessThan48HoursSinceLastWalking = col_factor(levels = c("0", "1")),
        IsLessThan60HoursSinceLastWalking = col_factor(levels = c("0", "1")),
        IsLessThan72HoursSinceLastWalking = col_factor(levels = c("0", "1")),
        IsRunning = col_factor(levels = c("0", "1")),
        IsRunningAtHawksRidge = col_factor(levels = c("0", "1")),
        IsRunningAtStringers = col_factor(levels = c("0", "1")),
        IsRunningAtRiverwalk = col_factor(levels = c("0", "1")),
        IsOtherRunning = col_factor(levels = c("0", "1")),
        IsLessThan12HoursSinceLastRunning = col_factor(levels = c("0", "1")),
        IsLessThan24HoursSinceLastRunning = col_factor(levels = c("0", "1")),
        IsLessThan36HoursSinceLastRunning = col_factor(levels = c("0", "1")),
        IsLessThan48HoursSinceLastRunning = col_factor(levels = c("0", "1")),
        IsLessThan60HoursSinceLastRunning = col_factor(levels = c("0", "1")),
        IsLessThan72HoursSinceLastRunning = col_factor(levels = c("0", "1")),
        IsHawksRidge = col_factor(levels = c("0", "1")),
        IsRiverwalk = col_factor(levels = c("0", "1")),
        IsStringers = col_factor(levels = c("0", "1")),
        IsWholeEnchilada = col_factor(levels = c("0", "1")),
        IsWhiteOakGreen = col_factor(levels = c("0", "1")),
        IsZone1 = col_factor(levels = c("0", "1")),
        IsLessThan12HoursSinceLastZone1 = col_factor(levels = c("0", "1")),
        IsLessThan24HoursSinceLastZone1 = col_factor(levels = c("0", "1")),
        IsLessThan36HoursSinceLastZone1 = col_factor(levels = c("0", "1")),
        IsLessThan48HoursSinceLastZone1 = col_factor(levels = c("0", "1")),
        IsLessThan60HoursSinceLastZone1 = col_factor(levels = c("0", "1")),
        IsLessThan72HoursSinceLastZone1 = col_factor(levels = c("0", "1")),
        IsZone2 = col_factor(levels = c("0", "1")),
        IsLessThan12HoursSinceLastZone2 = col_factor(levels = c("0", "1")),
        IsLessThan24HoursSinceLastZone2 = col_factor(levels = c("0", "1")),
        IsLessThan36HoursSinceLastZone2 = col_factor(levels = c("0", "1")),
        IsLessThan48HoursSinceLastZone2 = col_factor(levels = c("0", "1")),
        IsLessThan60HoursSinceLastZone2 = col_factor(levels = c("0", "1")),
        IsLessThan72HoursSinceLastZone2 = col_factor(levels = c("0", "1")),
        IsZone3 = col_factor(levels = c("0", "1")),
        IsLessThan12HoursSinceLastZone3 = col_factor(levels = c("0", "1")),
        IsLessThan24HoursSinceLastZone3 = col_factor(levels = c("0", "1")),
        IsLessThan36HoursSinceLastZone3 = col_factor(levels = c("0", "1")),
        IsLessThan48HoursSinceLastZone3 = col_factor(levels = c("0", "1")),
        IsLessThan60HoursSinceLastZone3 = col_factor(levels = c("0", "1")),
        IsLessThan72HoursSinceLastZone3 = col_factor(levels = c("0", "1")),
        IsZone4 = col_factor(levels = c("0", "1")),
        IsLessThan12HoursSinceLastZone4 = col_factor(levels = c("0", "1")),
        IsLessThan24HoursSinceLastZone4 = col_factor(levels = c("0", "1")),
        IsLessThan36HoursSinceLastZone4 = col_factor(levels = c("0", "1")),
        IsLessThan48HoursSinceLastZone4 = col_factor(levels = c("0", "1")),
        IsLessThan60HoursSinceLastZone4 = col_factor(levels = c("0", "1")),
        IsLessThan72HoursSinceLastZone4 = col_factor(levels = c("0", "1")),
        IsZone5 = col_factor(levels = c("0", "1")),
        IsLessThan12HoursSinceLastZone5 = col_factor(levels = c("0", "1")),
        IsLessThan24HoursSinceLastZone5 = col_factor(levels = c("0", "1")),
        IsLessThan36HoursSinceLastZone5 = col_factor(levels = c("0", "1")),
        IsLessThan48HoursSinceLastZone5 = col_factor(levels = c("0", "1")),
        IsLessThan60HoursSinceLastZone5 = col_factor(levels = c("0", "1")),
        IsLessThan72HoursSinceLastZone5 = col_factor(levels = c("0", "1")),
        IsLessThan12HoursSinceLastAnything = col_factor(levels = c("0", "1")),
        IsLessThan24HoursSinceLastAnything = col_factor(levels = c("0", "1")),
        IsLessThan36HoursSinceLastAnything = col_factor(levels = c("0", "1")),
        IsLessThan48HoursSinceLastAnything = col_factor(levels = c("0", "1")),
        IsLessThan60HoursSinceLastAnything = col_factor(levels = c("0", "1")),
        IsLessThan72HoursSinceLastAnything = col_factor(levels = c("0", "1"))
        ))
```

Let's view the data be executing the code block below. Alternatively, you can
double click `homes` in the *Environment* tab or issue the command
`View(homes)` from the console.

```{r}
#homes
HR_Exercise_Training
```

The following is a brief description of the columns:

- `sale_price` = price (in $) the home was sold for
- `above` = above ground square footage
- `basement` = basement square footage
- `walkout` = 0-1 variable indicating presence of walkout basement (base case = 0)
- `baths` = number of full baths
- `toilets` = number of half baths
- `beds` = number of bedrooms
- `condition` = condition of home, either Normal (base case), Below Normal, or Above Normal
- `age` = age of home (in years)

----

## Correlation Analysis

Before running a regression, it is often helpful to view a correlation matrix
for the variables in the data. To do this, we use the function `aa_cor(...)`
from the file `code_Linear_Regression_Review.R`, which has been loaded above.
Note that `aa_cor(...)` only returns pairwise correlations between non-factor
columns.


```{r}
#aa_cor(homes)
aa_cor(HR_Exercise_Training)
```

In this case, `above` has a very strong positive correlation with `sale_price`.
These two variables will be the basis of our first regression. Here is a
scatter plot of these variables.

```{r}
qplot(above, sale_price, data = homes)
qplot(DistanceInFeetPerMin, BPM, data = HR_Exercise_Training)
qplot(ElevationChangeInFeetPerMin, BPM, data = HR_Exercise_Training)
qplot(CaloriesPerMin, BPM, data = HR_Exercise_Training)
qplot(Temp, BPM, data = HR_Exercise_Training)
qplot(Humidity, BPM, data = HR_Exercise_Training)
qplot(DewPoint, BPM, data = HR_Exercise_Training)
qplot(BPM, Growth, data = HR_Exercise_Training)
```

----

## Simple Linear Regression

We are now ready to perform our first simple linear regression of `sale_price`
on `above`. This can be visualized with the following code, which plots a
scatter plot of the two variables along with their trendline (i.e., regression
line) and confidence interval for the regression.

```{r}
#qplot(above, sale_price, data = homes) + stat_smooth(method = "lm")
qplot(DistanceInFeetPerMin, BPM, data = HR_Exercise_Training) + stat_smooth(method = "lm")
qplot(ElevationChangeInFeetPerMin, BPM, data = HR_Exercise_Training) + stat_smooth(method = "lm")
qplot(CaloriesPerMin, BPM, data = HR_Exercise_Training) + stat_smooth(method = "lm")
qplot(Temp, BPM, data = HR_Exercise_Training) + stat_smooth(method = "lm")
qplot(Humidity, BPM, data = HR_Exercise_Training) + stat_smooth(method = "lm")
qplot(DewPoint, BPM, data = HR_Exercise_Training) + stat_smooth(method = "lm")
qplot(BPM, Growth, data = HR_Exercise_Training) + stat_smooth(method = "lm")
```

The `lm` in the code just above refers to the R function for running linear
regressions. We can call `lm(...)` directly, which is in fact necessary when we
want all the regression details. We save the results in an object which we call
`fit`. The code `sale_price ~ above` is called the regression "formula" and
means that `sale_price` is dependent and `above` is independent.

```{r}
fit <- lm(sale_price ~ above, data = homes)
fit1 <- lm(Growth ~ Exercise, data = HR_Exercise_Training)
fit2 <- lm(Growth ~ ExerciseValue, data = HR_Exercise_Training)
fit2 <- lm(Growth ~ TimeInMin, data = HR_Exercise_Training)
fit3 <- lm(Growth ~ Course, data = HR_Exercise_Training)
fit4 <- lm(Growth ~ CourseValue, data = HR_Exercise_Training)
fit4 <- lm(Growth ~ BPM, data = HR_Exercise_Training)
fit5 <- lm(Growth ~ MaxBPM, data = HR_Exercise_Training)
fit6 <- lm(Growth ~ Calories, data = HR_Exercise_Training)
#fit7 <- lm(Growth ~ Zone, data = HR_Exercise_Training)
#fit8 <- lm(Growth ~ Miles, data = HR_Exercise_Training)
#fit9 <- lm(Growth ~ MilesPerHour, data = HR_Exercise_Training)
fit10 <- lm(Growth ~ Temp, data = HR_Exercise_Training)
fit11 <- lm(Growth ~ Humidity, data = HR_Exercise_Training)
fit12 <- lm(Growth ~ DewPoint, data = HR_Exercise_Training)
#fit13 <- lm(Growth ~ ElevationChangeInFeet, data = HR_Exercise_Training)
#fit14 <- lm(Growth ~ ElevationChangeInFeetPerFoot, data = HR_Exercise_Training)
#fit15 <- lm(Growth ~ ElevationChangeInFeetPerMin, data = HR_Exercise_Training)
fit16 <- lm(Growth ~ CaloriesPerMin, data = HR_Exercise_Training)
fit17 <- lm(Growth ~ CaloriesPerFoot, data = HR_Exercise_Training)
#fit18 <- lm(Growth ~ IsYoga, data = HR_Exercise_Training)
#fit19 <- lm(Growth ~ IsRunning, data = HR_Exercise_Training)
#fit20 <- lm(Growth ~ IsWalking, data = HR_Exercise_Training)
#fit21 <- lm(Growth ~ IsBiking, data = HR_Exercise_Training)
```

Once we have the regression model in `fit`, it is important to critique the
model to see if it's a good model. We run the function `aa_critique_fit(...)`,
which comes from `code_Linear_Regression_Review.R`, on `fit` and save its
results in an object that we call `crit`.

```{r}
crit <- aa_critique_fit(fit)
#crit1 <- aa_critique_fit(fit1)
```

Inside `crit` are several objects that will help us understand and critique the
fit. First, the object `crit$formula` shows the fitted regression formula.
This is "what we tell the boss" and is, in some sense, the most important
feature of the regression model.

```{r}
print(crit$formula)
#print(crit1$formula)
```

Next, the object `crit$summary` contains a full summary of the regression
model, including the standard error, adjusted $R^2$, and p-value associated
with the independent variable `above`.

```{r}
crit$summary
#crit1$summary
```

We can also view 95% confidence intervals on the intercept and coefficients by
examining `crit$confint`.

```{r}
crit$confint
#crit1$confint
```

The adjusted $R^2$, standard error, and average $Y$ values are also saved in
`crit`:

```{r}
crit$R2
#crit1$R2
```

```{r}
crit$Se
#crit1$Se
```

```{r}
crit$mean_Y
crit1$mean_Y
```

The plot of residuals versus the fitted values is contained in
`crit$residual_plot`, and the residual histogram is contained in
`crit$residual_histogram`.

```{r}
crit$residual_plot
#crit1$residual_plot
```

```{r}
crit$residual_histogram
#crit1$residual_histogram
```

### Is this a good model?

Here is a summary of some key aspects of this model:

- The variable `above` is significant in the regression because its
  p-value is below 0.05.
- The coefficient of `above` makes sense as we would expect home
  values to increase with more above ground square footage.
- Notice that the intercept confidence interval contains 0, which
  would be the expected price for a home with `above = 0`. This
  makes sense.
- The adjusted $R^2$ value `r crit$R2` is a medium-level value.
- The standard error value is about 30% of the average sale_price, which
  means that the regression misses the data by about 30% on average. This
  is fairly high.
- The residual histogram is fairly normal, which is good.
- The residual plot does not look like a "strip of random points." Rather
  it has more of a cone shape.

In summary, the model looks pretty good except that the standard error is a bit
high and the residual plot doesn't have the desired shape.

----

## Adding Another Independent Variable

To expand the model to include more independent variables, we simply expand the
formula `sale_price ~ above` to `sale_price ~ above + walkout`.  The key is to
use the `+` character to add new variables into the model.  In this case, we
are adding a factor variable, `walkout`. Recall that 0 is the base case of
`walkout`.


```{r}
#fit <- lm(sale_price ~ above + walkout, data = homes)
fit_some <- lm(Growth ~ ExerciseValue +
+ IsHawksRidge + IsRiverwalk + IsWholeEnchilada + IsWhiteOakGreen + IsStringers
+ Temp + DewPoint + Humidity
, data = HR_Exercise_Training, na.action=na.omit)
```

As before, we can run `aa_critique_fit(...)` to provide output that needed to
critique the model.

```{r}
#crit <- aa_critique_fit(fit)
crit_some <- aa_critique_fit(fit_some)
```

In this case, let us examine the formula.


```{r}
#print(crit$formula)
print(crit_some$formula)
```

The term `walkout1` in this formula should be interpreted as follows: when the
home has a walkout basement (`walkout = 1`), then the home value is adjusted by
the specified coefficient; otherwise, if the home does not have a walkout
basement (`walkout = 0`), then there is no adjustment to the home value.  A
similar meaning is shown in the regression summary.

```{r}
#crit$summary
crit_some$summary
```

----

## The Full Regression

To complete this example, we now add in all of the independent variables and
critique the model:

```{r}
fit <- lm(sale_price ~ above + basement + walkout + baths + toilets + beds +
    condition + age, data = homes)
```

```{r}
#crit <- aa_critique_fit(fit)
crit_some <- aa_critique_fit(fit_some)
```

```{r}
#print(crit$formula)
print(crit_some$formula)
```

```{r}
#crit$summary
crit_some$summary
```

```{r}
crit$residual_plot
crit_some$residual_plot
```

```{r}
crit$residual_histogram
crit_some$residual_histogram
```


A look at how large the mistakes are:

```{r}
crit_some$Se / crit_some$mean_Y
```



A few points are in order:

- Notice that coefficients are given for the levels `Below_Normal`
  and `Above_Normal` of `condition`. This is because `Normal` is the
  base case.
- The variables `toilets` is insignificant in this regression. It
  could be removed in order to keep the model simpler, but we will
  leave it in.
- Compared to our previous regressions, we see better standard error
  and adjusted $R^2$. However, the cone shape of the residual plot
  still remains.

Finally, we check the VIFs to verify that multicollinearity is not present and
possibly causing problems. Recall any VIF above 5.0 is considered evidence of
multicollinearity.

```{r}
#crit$vif
crit_some$vif
```
