---
title: "Time Series: Example 1"
author: "Samuel Burer (samuel-burer@uiowa.edu)"
date: "Updated: `r format(Sys.time(), '%B %e, %Y')`"
output: html_notebook
---

----

# Preparation

If you are following along in RStudio, to prepare for this example,
choose the RStudio menu item *Session > Restart and Clear Output*.
Alternatively, you can choose *Run > Restart R and Clear Output* in the
notebook code window. You will also need to install the `ggplot2` and
`forecast` packages and to source `code_Time_Series.R`.

```{r}
# install.packages(c("ggplot2", "forecast")) # Paste into console to install
source("code_Time_Series.R")
```

# Data and Getting Started

The file `data_Time_Series.Rdata` contains several time series that
will be the basis of our example. In particular, we will work only with
univariate time series, i.e., time series that contain just one column
of data. Let's load the data:

```{r}
load("data_Time_Series.Rdata")
library(readr)
HRExercise <- read_csv("HRExercise.csv", 
    col_types = cols(Date = col_date(format = "%m/%d/%y"), 
        Time = col_time(format = "%T"), DateTime = col_datetime(format = "%m/%d/%y %T"), 
        Exercise = col_factor(levels = c("Yoga", "Walking", "Running", "Biking", "Trail Walking", "Trail Running", "Mountain Biking")), 
        ExerciseValue = col_factor(levels = c("0", "1", "2", "3", "4", "5", "6")), 
        Course = col_factor(levels = c("Alta Vista", 
            "Deer Hollow", "Enterprise South", 
            "Four Pass Loop", "Full Practice", 
            "Greenway", "Hawk's Ridge", "Montreat", 
            "Montreat Trails", "Restorative", 
            "Riverwalk", "Southern Soul", 
            "Stringers", "White Oak - Green", 
            "White Sands Cove", "Whole Enchilada")), 
        CourseValue = col_factor(levels = c("0", "1", "2", "3")),
        Zone = col_factor(levels = c("1", "2", "3", "4", "5")),
        IsYoga = col_factor(levels = c("0", "1")), 
        IsFullPractice = col_factor(levels = c("0", "1")), 
        IsSouthernSoul = col_factor(levels = c("0", "1")), 
        IsOtherYoga = col_factor(levels = c("0", "1")), 
        IsLessThan12HoursSinceLastYoga = col_factor(levels = c("0", "1")), 
        IsLessThan24HoursSinceLastYoga = col_factor(levels = c("0", "1")), 
        IsLessThan36HoursSinceLastYoga = col_factor(levels = c("0", "1")), 
        IsLessThan48HoursSinceLastYoga = col_factor(levels = c("0", "1")), 
        IsLessThan60HoursSinceLastYoga = col_factor(levels = c("0", "1")), 
        IsLessThan72HoursSinceLastYoga = col_factor(levels = c("0", "1")), 
        IsBiking = col_factor(levels = c("0", "1")), 
        IsBikingAtWhiteOakGreen = col_factor(levels = c("0", "1")), 
        IsBikingAtWholeEnchilada = col_factor(levels = c("0", "1")), 
        IsBikingAtStringers = col_factor(levels = c("0", "1")), 
        IsBikingAtRiverwalk = col_factor(levels = c("0", "1")),
        IsBikingAtGreenway = col_factor(levels = c("0", "1")),
        IsOtherBiking = col_factor(levels = c("0", "1")),
        IsLessThan12HoursSinceLastBiking = col_factor(levels = c("0", "1")),
        IsLessThan24HoursSinceLastBiking = col_factor(levels = c("0", "1")),
        IsLessThan36HoursSinceLastBiking = col_factor(levels = c("0", "1")),
        IsLessThan48HoursSinceLastBiking = col_factor(levels = c("0", "1")),
        IsLessThan60HoursSinceLastBiking = col_factor(levels = c("0", "1")),
        IsLessThan72HoursSinceLastBiking = col_factor(levels = c("0", "1")),
        IsWalking = col_factor(levels = c("0", "1")),
        IsWalkingAtHawksRidge = col_factor(levels = c("0", "1")),
        IsWalkingAtFourPassLoop = col_factor(levels = c("0", "1")),
        IsWalkingAtRiverwalk = col_factor(levels = c("0", "1")),
        IsOtherWalking = col_factor(levels = c("0", "1")),
        IsLessThan12HoursSinceLastWalking = col_factor(levels = c("0", "1")),
        IsLessThan24HoursSinceLastWalking = col_factor(levels = c("0", "1")),
        IsLessThan36HoursSinceLastWalking = col_factor(levels = c("0", "1")),
        IsLessThan48HoursSinceLastWalking = col_factor(levels = c("0", "1")),
        IsLessThan60HoursSinceLastWalking = col_factor(levels = c("0", "1")),
        IsLessThan72HoursSinceLastWalking = col_factor(levels = c("0", "1")),
        IsRunning = col_factor(levels = c("0", "1")),
        IsRunningAtHawksRidge = col_factor(levels = c("0", "1")),
        IsRunningAtStringers = col_factor(levels = c("0", "1")),
        IsRunningAtRiverwalk = col_factor(levels = c("0", "1")),
        IsOtherRunning = col_factor(levels = c("0", "1")),
        IsLessThan12HoursSinceLastRunning = col_factor(levels = c("0", "1")),
        IsLessThan24HoursSinceLastRunning = col_factor(levels = c("0", "1")),
        IsLessThan36HoursSinceLastRunning = col_factor(levels = c("0", "1")),
        IsLessThan48HoursSinceLastRunning = col_factor(levels = c("0", "1")),
        IsLessThan60HoursSinceLastRunning = col_factor(levels = c("0", "1")),
        IsLessThan72HoursSinceLastRunning = col_factor(levels = c("0", "1")),
        IsHawksRidge = col_factor(levels = c("0", "1")),
        IsRiverwalk = col_factor(levels = c("0", "1")),
        IsWholeEnchilada = col_factor(levels = c("0", "1")),
        IsWhiteOakGreen = col_factor(levels = c("0", "1")),
        IsStringers = col_factor(levels = c("0", "1")),
        IsZone1 = col_factor(levels = c("0", "1")),
        IsLessThan12HoursSinceLastZone1 = col_factor(levels = c("0", "1")),
        IsLessThan24HoursSinceLastZone1 = col_factor(levels = c("0", "1")),
        IsLessThan36HoursSinceLastZone1 = col_factor(levels = c("0", "1")),
        IsLessThan48HoursSinceLastZone1 = col_factor(levels = c("0", "1")),
        IsLessThan60HoursSinceLastZone1 = col_factor(levels = c("0", "1")),
        IsLessThan72HoursSinceLastZone1 = col_factor(levels = c("0", "1")),
        IsZone2 = col_factor(levels = c("0", "1")),
        IsLessThan12HoursSinceLastZone2 = col_factor(levels = c("0", "1")),
        IsLessThan24HoursSinceLastZone2 = col_factor(levels = c("0", "1")),
        IsLessThan36HoursSinceLastZone2 = col_factor(levels = c("0", "1")),
        IsLessThan48HoursSinceLastZone2 = col_factor(levels = c("0", "1")),
        IsLessThan60HoursSinceLastZone2 = col_factor(levels = c("0", "1")),
        IsLessThan72HoursSinceLastZone2 = col_factor(levels = c("0", "1")),
        IsZone3 = col_factor(levels = c("0", "1")),
        IsLessThan12HoursSinceLastZone3 = col_factor(levels = c("0", "1")),
        IsLessThan24HoursSinceLastZone3 = col_factor(levels = c("0", "1")),
        IsLessThan36HoursSinceLastZone3 = col_factor(levels = c("0", "1")),
        IsLessThan48HoursSinceLastZone3 = col_factor(levels = c("0", "1")),
        IsLessThan60HoursSinceLastZone3 = col_factor(levels = c("0", "1")),
        IsLessThan72HoursSinceLastZone3 = col_factor(levels = c("0", "1")),
        IsZone4 = col_factor(levels = c("0", "1")),
        IsLessThan12HoursSinceLastZone4 = col_factor(levels = c("0", "1")),
        IsLessThan24HoursSinceLastZone4 = col_factor(levels = c("0", "1")),
        IsLessThan36HoursSinceLastZone4 = col_factor(levels = c("0", "1")),
        IsLessThan48HoursSinceLastZone4 = col_factor(levels = c("0", "1")),
        IsLessThan60HoursSinceLastZone4 = col_factor(levels = c("0", "1")),
        IsLessThan72HoursSinceLastZone4 = col_factor(levels = c("0", "1")),
        IsZone5 = col_factor(levels = c("0", "1")),
        IsLessThan12HoursSinceLastZone5 = col_factor(levels = c("0", "1")),
        IsLessThan24HoursSinceLastZone5 = col_factor(levels = c("0", "1")),
        IsLessThan36HoursSinceLastZone5 = col_factor(levels = c("0", "1")),
        IsLessThan48HoursSinceLastZone5 = col_factor(levels = c("0", "1")),
        IsLessThan60HoursSinceLastZone5 = col_factor(levels = c("0", "1")),
        IsLessThan72HoursSinceLastZone5 = col_factor(levels = c("0", "1")),
        IsLessThan12HoursSinceLastAnything = col_factor(levels = c("0", "1")),
        IsLessThan24HoursSinceLastAnything = col_factor(levels = c("0", "1")),
        IsLessThan36HoursSinceLastAnything = col_factor(levels = c("0", "1")),
        IsLessThan48HoursSinceLastAnything = col_factor(levels = c("0", "1")),
        IsLessThan60HoursSinceLastAnything = col_factor(levels = c("0", "1")),
        IsLessThan72HoursSinceLastAnything = col_factor(levels = c("0", "1"))
        ))
```

The data sets are as follows:

- `cpi` = annual Consumer Price Index since 1947, which is indexed to
  the value 1.0
- `elec` = monthly net electricity generation in the U.S. (in kWh)
- `google` = daily closing price (in $) of Google stock (start date =
  February 25, 2013)
- `lettuce` = monthly Producer Price Index for lettuce since 1947; the
  average of 1982 is indexed to 100.0
- `qc` = quality control data; 5 measurements per day, for 25
  consecutive days; measurements are taken at 6:00, 10:00, 14:00, 18:00,
  and 22:00
- `rain` = annual rainfall (in inches) in Philadelphia
- `soda` = monthly soda sales (in thousands of cases) for a soda
  distributor
- `ui` = total Fall plus Spring semester enrollment at the University
  of Iowa

In the code chunk below, you can uncomment whichever series you would
like to work with in this example. Then the time series will be
referenced in the object called `myts` throughout this example.

```{r}
# myts <- cpi
# myts <- elec
myts <- google
# myts <- lettuce
# myts <- qc
# myts <- rain
# myts <- soda
# myts <- ui

efficiency <- data.frame(HRExercise$Efficiency)
efficiency <- ts(data = efficiency, start = c(as.Date(HRExercise$DateTime[1])), end=c(as.Date(HRExercise$DateTime[113])), frequency = 1)
```

Let's first look at the raw data:

```{r}
myts
efficiency
```

The command `autoplot(...)` can be used to provide a quick plot of the
time series.

```{r}
#autoplot(myts)
autoplot(efficiency)
```

If the time series exhibits unstable variation (or unstable fluctuation)
over time, you should log the time series. You should also log the time
series if you suspect an exponential trend over time. If so, you can
uncomment the following two lines to resave `myts` in logged form and
print a new plot.

```{r}
#myts <- log(myts)
#autoplot(myts)

efficiency <- log(efficiency)
autoplot(efficiency)
```

If you do decide to log your time series, then keep this in mind for the
rest of the example. Any analysis will be in the log, and you'll need to
do exp to get the analysis back in the original data.

# Exploration

Before doing any sophisticated analysis of a time series, it is
important to get a basic feel for the data. One of the best ways is
to look at the decomposition of `myts`, which breaks the time series
down into a trend component, a seasonal component (as applicable),
and a remainder component. Pay attention to the $y$ scales in the
decomposition plot!

```{r}
#decomp <- mstl(myts)
#autoplot(decomp)

decomp_efficiency <- mstl(efficiency)
autoplot(decomp_efficiency)
```

It's also helpful to know if the series has any autocorrelation issues,
which can be viewed using the `Acf(...)` function. Note the capital
letter "A" but lowercase letters "cf". Autocorrelation can also reveal
any seasonality patterns---which you may or may not already be aware of.

```{r}
Acf(myts)
```
