---
title: "HRExercise"
output: html_document
---

## Preparation

Ensure all relevant packages are installed.

```{r}
#install.packages(c("rmarkdown", "car", "ggplot2", "forecast", "dplyr")) 
#install.packages("stats")
```


Load libraries and supporting functions.

```{r}
library(readr)
library(dplyr)
source("code_Linear_Regression_Review.R")
source("code_Time_Series.R")
source("HRExercise-Data.R")
```


----

## Correlation Analysis

Uses pre-built, loaded function `aa_cor()`

```{r}
aa_cor(HRExercise)
```


Scatter plot for highly correlated variables.

```{r}
qplot(Efficiency, BPM, data = HRExercise)
qplot(Efficiency, CaloriesPerMin, data = HRExercise)
qplot(Efficiency, Work, data = HRExercise)
```


The command `autoplot(...)` can be used to provide a quick plot of the
time series.

```{r}
autoplot(efficiency)
autoplot(efficiency_lag)
autoplot(bpm)
autoplot(caloriespermin)
autoplot(calories)
autoplot(timeinmin)
autoplot(elevationchangeinfeetpermin)
autoplot(elevationchangeinfeet)
autoplot(distanceinfeetpermin)
autoplot(distanceinfeet)
autoplot(temp)
autoplot(dewpoint)
autoplot(exercise)
autoplot(course)
autoplot(islessthan24hourssincelastanything)
autoplot(islessthan36hourssincelastanything)
autoplot(islessthan48hourssincelastanything)
```


If the time series exhibits unstable variation (or unstable fluctuation)
over time, you should log the time series. You should also log the time
series if you suspect an exponential trend over time. 

```{r}
efficiency_log <- log(efficiency)
autoplot(efficiency_log)

BPM_log <- log(BPM)
autoplot(BPM_log)

CaloriesPerMin_log <- log(CaloriesPerMin)
autoplot(CaloriesPerMin_log)
```


----

## Decomposition

A look at the decomposition, which breaks the time series
down into a trend component, a seasonal component (as applicable),
and a remainder component.

```{r}
efficiency_decomp <- mstl(efficiency)
autoplot(efficiency_decomp)

bpm_decomp <- mstl(bpm)
autoplot(bpm_decomp)

caloriespermin_decomp <- mstl(caloriespermin)
autoplot(caloriespermin_decomp)

calories_decomp <- mstl(calories)
autoplot(calories_decomp)

timeinmin_decomp <- mstl(timeinmin)
autoplot(timeinmin_decomp)

distanceinfeetpermin_decomp <- mstl(distanceinfeetpermin)
autoplot(distanceinfeetpermin_decomp)

distanceinfeet_decomp <- mstl(distanceinfeet)
autoplot(distanceinfeet_decomp)

elevationchangeinfeetpermin_decomp <- mstl(elevationchangeinfeetpermin)
autoplot(elevationchangeinfeetpermin_decomp)

elevationchangeinfeet_decomp <- mstl(elevationchangeinfeet)
autoplot(elevationchangeinfeet_decomp)

temp_decomp <- mstl(temp)
autoplot(temp_decomp)

dewpoint_decomp <- mstl(dewpoint)
autoplot(dewpoint_decomp)

exercise_decomp <- mstl(exercise)
autoplot(exercise_decomp)

course_decomp <- mstl(course)
autoplot(course_decomp)

islessthan24hourssincelastanything_decomp <- mstl(islessthan24hourssincelastanything)
autoplot(islessthan24hourssincelastanything_decomp)

islessthan36hourssincelastanything_decomp <- mstl(islessthan36hourssincelastanything)
autoplot(islessthan36hourssincelastanything_decomp)

islessthan48hourssincelastanything_decomp <- mstl(islessthan48hourssincelastanything)
autoplot(islessthan48hourssincelastanything_decomp)
```


----

## Autocorrelation

Any autocorrelation issues can be viewed using the `Acf()` function. 

```{r}
Acf(efficiency)
Acf(bpm)
Acf(caloriespermin)
Acf(calories)
Acf(timeinmin)
Acf(distanceinfeetpermin)
Acf(distanceinfeet)
Acf(elevationchangeinfeetpermin)
Acf(elevationchangeinfeet)
Acf(temp)
Acf(dewpoint)
Acf(exercise)
Acf(course)
Acf(islessthan24hourssincelastanything)
Acf(islessthan36hourssincelastanything)
Acf(islessthan48hourssincelastanything)

```


----

## Linear Regression 

First, plot a scatter plot of any two variables along with their trendline (i.e., regression
line) and confidence interval for the regression.

```{r}
qplot(Efficiency, BPM, data = HRExercise) + stat_smooth(method = "lm")
qplot(Efficiency, CaloriesPerMin, data = HRExercise) + stat_smooth(method = "lm")
```


Now, run linear regression, adding multiple variables, storing in fit.

# No Time Series

```{r}
fit <- lm(Efficiency ~ CaloriesPerMin 
+ TimeInMin 
+ BPM
#+ DistanceInFeetPerMin
+ Temp + DewPoint 
, data = HRExercise, na.action=na.omit)
```


Run pre-built, loaded function `aa_critique_fit()` to provide output that needed to
critique the model.

```{r}
crit <- aa_critique_fit(fit)
```


The object `crit$formula` shows the fitted regression formula.

```{r}
print(crit$formula)
```


The object `crit$summary` contains a full summary of the regression
model, including the standard error, adjusted $R^2$, and p-value associated
with the independent variables.

```{r}
crit$summary
```

View 95% confidence intervals on the intercept and coefficients.

```{r}
crit$confint
```

The adjusted $R^2$, standard error, and average $Y$ values.

```{r}
crit$R2
crit$Se
crit$mean_Y
```

The plot of residuals versus the fitted values is contained in
`crit$residual_plot`, and the residual histogram is contained in
`crit$residual_histogram`.

```{r}
crit$residual_plot
crit$residual_histogram
```

A look at how large the mistakes are:

```{r}
crit$Se / crit$mean_Y
```


Check the VIFs to verify that multicollinearity is not present and
possibly causing problems.

```{r}
crit$vif
```


# Time Series

```{r}
#fit <- tslm(efficiency ~ efficiency_lag)
fit <- tslm(efficiency ~ 
+ timeinmin #+ timeinmin_lag
+ elevationchangeinfeet + elevationchangeinfeet_lag
#+ elevationchangeinfeetpermin #+ elevationchangeinfeetpermin_lag
+ distanceinfeet + distanceinfeet_lag
+ distanceinfeetpermin #+ distanceinfeetpermin_lag
+ temp + dewpoint #+ humidity
#+ exercise
#+ course
#+ islessthan12hourssincelastanything #+ islessthan24hourssincelastanything + islessthan36hourssincelastanything 
#+ time
)
summary(fit)
```

```{r}

fit <- tslm(efficiency ~ 
+ timeinmin 
+ distanceinfeet
#+ distanceinfeetpermin
+ elevationchangeinfeet
+ elevationchangeinfeetpermin
+ temp + dewpoint #+ humidity
#+ exercise
+ isbiking +iswalking #+ isrunning 
#+ isyoga
+ course
#+ southernsoul 
#+ wholeenchilada + riverwalk + hawksridge + stringers + whiteoak
#+ islessthan12hourssincelastanything + islessthan24hourssincelastanything #+ islessthan36hourssincelastanything 

#+ islessthan24hourssincelastrunning #+ islessthan36hourssincelastrunning
#+ islessthan24hourssincelastbiking #+ islessthan36hourssincelastbiking
#+ islessthan48hourssincelastwalking #+ islessthan36hourssincelastwalking
)
summary(fit)
```


Fitted values plotted along with the data itself:

```{r}
aa_plot_fitted(fit)
```


Calculate the total accuracy of the regression using several
measures.

```{r}
accuracy(fit)
```


Check the normality and autocorrelation of the residual
errors.

```{r}
checkresiduals(fit)
```



View the forecasts with 80% and 95% confidence bands.

```{r}
forecast(fit)
```



























